---
title: "HR Analytics Hackathon"
output: html_notebook
---






```{r}
# Import datasets
require(readr)
train <- read_csv("E:/coaching/Data Science/R/dataset/HR Analytics/dataset/train_LZdllcl.csv")
test <-  read_csv("E:/coaching/Data Science/R/dataset/HR Analytics/dataset/test_2umaH9m.csv")

# Get a feel of the data at hand
head(train)

```



```{r}

# Combine datasets for cleaning
require(dplyr)
master <- bind_rows(train, test)

str(master)

```



```{r}
#### ---- Data cleaning ---- ####

# Avoid case mismatch possibility
master <- mutate_if(master, is.character, tolower)

```



```{r}
# Check duplicates

master <- distinct(master)

```



```{r}
# Check missing values
colSums(is.na(master))

```



```{r}
3443 / nrow(master) * 100

5936 / nrow(master) * 100

```



```{r}
# NA's in previous year rating
summary(as.factor(master$previous_year_rating))

# Check where previous_year_rating is NA
df <- master[which(is.na(master$previous_year_rating)), ]

View(df)
```



```{r}
# We see that where previous_year_rating is NA, length_of_service is "1".
# So NA's seem justified.
require(ggplot2)

ggplot(master[1:54808, ], aes(x = previous_year_rating, fill = as.factor(is_promoted))) + 
  geom_bar(position = 'fill')

```
And it is significant!



```{r}
master$previous_year_rating[which(is.na(master$previous_year_rating))] <- 'freshers'

ggplot(master[1:54808, ], aes(x = previous_year_rating, fill = as.factor(is_promoted))) + 
  geom_bar(position = 'fill')

```



```{r}
# Blanks in education
summary(as.factor(master$education))

ggplot(master[1:54808,], aes(x = education, fill = as.factor(is_promoted))) + geom_bar(position = 'fill')

```



```{r}
edu_gp <- group_by(train, education) %>% summarise(prom_perc = round(sum(is_promoted) / n() * 100, 2))

edu_gp

```



```{r}
# Lets keep it simple for now.
master$education[which(is.na(master$education))] <- "unknown"

table(master$education)

```



```{r}
# Univariate / Bivariate Analysis ----

# 1 employee_id
n_distinct(master$employee_id) == nrow(master)

```



```{r}
# 2 department

ggplot(master, aes(x = department)) + 
  geom_bar(fill = 'skyblue',color = 'black') + coord_flip()

dept_df <- group_by(master, department) %>% 
  summarise(dept_influence = round(sum(is_promoted, na.rm = T)/n()*100, 2)) %>%
  arrange(dept_influence)

dept_df

```

Is there departmental bias ???





```{r}

# 3 region

ggplot(master, aes(x = region)) + 
  geom_bar(fill = 'skyblue',color = 'black') + coord_flip()

reg_df <- group_by(master, region) %>% 
          summarise(reg_strength = n()) %>%
          arrange(reg_strength)

# Way too many categories in region. Pattern detection not possible.

```



```{r}
master$region <- NULL

```



```{r}

# 5 gender
ggplot(master[!is.na(master$is_promoted),],aes(x = gender, fill =  gender)) + geom_bar() + coord_polar()


ggplot(master[!is.na(master$is_promoted),], 
       aes(x = gender, fill =  as.factor(is_promoted))) + 
  geom_bar(position = 'fill')
# Lesser female employees overall. But no apparent gender bias in promotions.


```



```{r}
# 6 recruitment_channel

channel_df <- group_by(master, recruitment_channel) %>% 
  summarise(channel_influence = round(sum(is_promoted, na.rm = T)/n()*100,2)) %>%
  arrange(channel_influence)

channel_df
# Clearly, referred employees outdo others.


```



```{r}

# 7 no_of_trainings

# Check outliers
plot(quantile(master$no_of_trainings, seq(0,1,0.01)))
quantile(master$no_of_trainings, seq(0,1,0.01))

master$no_of_trainings[master$no_of_trainings > 4] <- 4


```



```{r}
ggplot(master[!is.na(master$is_promoted),], aes(x = no_of_trainings, fill = as.factor(is_promoted))) +
  geom_bar(position = 'fill')

# More the no of trainings required, lesser the chance of promotion.

```




```{r}

# 8 Age

# Check outliers
plot(quantile(master$age, seq(0,1,0.01)))
# No outliers

ggplot(train, aes(x = age, fill = as.factor(is_promoted))) + 
  geom_histogram(binwidth = 10, color = 'black', position = 'fill')

```



```{r}

# 10 length_of_service

# Check outliers
plot(quantile(master$length_of_service, seq(0,1,0.01)))
quantile(master$length_of_service, seq(0,1,0.01))

master$length_of_service[master$length_of_service > 20] <- 20


```



```{r}
ggplot(master[!is.na(master$is_promoted),], 
       aes(x = length_of_service, 
           fill = as.factor(is_promoted))) + 
  geom_histogram(binwidth = 5, position = 'fill',color = 'black')


```



```{r}


# 11 KPIs_met>80%

ggplot(master[!is.na(master$is_promoted),], 
       aes(x = `KPIs_met >80%`, fill = as.factor(is_promoted))) + 
  geom_bar(position = 'fill',color = 'black')

# Clearly, meeting KPI matters for promotion


```



```{r}

# 12 awards_won?

ggplot(master[!is.na(master$is_promoted),], 
       aes(x = `awards_won?`, fill = as.factor(is_promoted))) + 
  geom_bar(position = 'fill',color = 'black')

# It highly impacts chances of promotion.



```



```{r}
# 13 avg_training_score
plot(quantile(master$avg_training_score, seq(0,1,0.01)))
quantile(master$avg_training_score, seq(0,1,0.01))

master$avg_training_score[master$avg_training_score > 91] <- 91
master$avg_training_score[master$avg_training_score < 44] <- 44

```


```{r}


ggplot(master[!is.na(master$is_promoted),], 
       aes(x = avg_training_score, 
           fill = as.factor(is_promoted))) + 
  geom_histogram(binwidth = 10, position = 'fill',color = 'black')



```



Feature Engineering ----

```{r}
# Creating some new metrics which may help in prediction.

# total training score of employees
master$tot_training_score <- master$no_of_trainings * master$avg_training_score

# Age of employee when joined company
master$start_time_age <- master$age - master$length_of_service

```



```{r}

# Lets normalize continuous variables.
master[ , c(6,7,9,12,14,15)] <- sapply(master[ , c(6,7,9,12,14,15)], scale)

glimpse(master) 

# EDA Complete...

```





```{r}
require(dummies)
master <- dummy.data.frame(as.data.frame(master))


train <- master[which(!is.na(master$is_promoted)),]

test.data <- master[which(is.na(master$is_promoted)),]


require(caTools)
set.seed(999)
index <- sample.split(train$is_promoted, SplitRatio = 0.75)

trn.data <- train[index, ]
val.data <- train[!index, ]

```

EDA Complete !



Model Building ----

```{r}

rm(channel_df, dept_df, df, edu_gp, master, reg_df, test, train)

View(trn.data)

```



1. Logistic Regression ----

```{r}
model_1 <- glm(is_promoted ~ ., data = trn.data[,-1], family = 'binomial')

summary(model_1)


```



Step-wise reduction

```{r}
model_2 <- step(model_1)

summary(model_2)

```


```{r}
require(car)

sort(vif(model_2))

```



removing dedpartmentoperations. High VIF and lesser impact on AIC.


```{r}
model_3 <- glm(formula = is_promoted ~ departmentanalytics + departmentfinance + 
    departmenthr + departmentlegal + departmentprocurement + 
    `departmentr&d` + `departmentsales & marketing` + `educationbachelor's` + 
    `educationmaster's & above` + no_of_trainings + age + previous_year_rating1 + 
    previous_year_rating2 + previous_year_rating4 + previous_year_rating5 + 
    length_of_service + `KPIs_met >80%` + `awards_won?` + avg_training_score, 
    family = "binomial", data = trn.data[, -1])


summary(model_3)

```



```{r}
sort(vif(model_3))
```






```{r}
round(prop.table(table(trn.data$is_promoted)) * 100, 2)
# Target class has huge imbalance.


```

Predicting.

```{r}
prob_prom <- predict(model_3, newdata = val.data, type = 'response')

summary(prob_prom)

pred_prom <- as.factor(ifelse(prob_prom > 0.085, 1, 0))

act_prom <- as.factor(val.data$is_promoted)

require(caret)
confusionMatrix(pred_prom, act_prom, positive = '1')

```



```{r}
prob_prom <- predict(model_3, newdata = test.data, type = 'response')

test.data$is_promoted <- ifelse(prob_prom > 0.085, 1, 0)

write.csv(test.data[, c('employee_id', 'is_promoted')], 'sub_log.csv', row.names = F)

```










2. Decision Tree ----




```{r}
trn.data$is_promoted <- as.factor(trn.data$is_promoted)
val.data$is_promoted <- as.factor(val.data$is_promoted)

require(rpart)
dt_model <- rpart(is_promoted ~ ., data = trn.data[,-1])

require(rpart.plot)
prp(dt_model)

```

Underfitted tree due to class imbalance.



```{r}

table(trn.data$is_promoted[trn.data$avg_training_score < 2])

table(trn.data$is_promoted[trn.data$avg_training_score >= 2])

```



Let's predict

```{r}

dt_pred <- predict(dt_model, newdata = val.data)
summary(dt_pred)

dt_prom <- as.factor(ifelse(dt_pred[,2] > 0.077, "1", "0"))

confusionMatrix(dt_prom, act_prom, positive = '1')

```

Poor validation. Not worthy to go further.



3. K-Nearest Neighbors ----

```{r}
require(class)

knn_prom <- knn(train = select(trn.data, -c(employee_id, is_promoted)), 
                cl = trn.data$is_promoted, 
                test = select(val.data, -c(employee_id, is_promoted)), 
                k = 5)

confusionMatrix(knn_prom, act_prom, positive = '1')

```

K-NN not working due to class imbalance in target variable.




```{r}
names(trn.data)
```

4. Random Forest

```{r}
require(randomForest)

names(trn.data)[c(8,9,11,12,13,29,30)] <- c("department_r_d", "department_sales_marketing",
                                            "education_bachelors", "education_below_secondary",
                                            "education_masters", "KPIs_met_80",
                                            "awards_won")

names(val.data)[c(8,9,11,12,13,29,30)] <- c("department_r_d", "department_sales_marketing",
                                            "education_bachelors", "education_below_secondary",
                                            "education_masters", "KPIs_met_80",
                                            "awards_won")

names(test.data)[c(8,9,11,12,13,29,30)] <- c("department_r_d", "department_sales_marketing",
                                             "education_bachelors", "education_below_secondary",
                                             "education_masters", "KPIs_met_80",
                                             "awards_won")

rm(dt_model, dt_pred, model_1, model_2, dt_prom, index, knn_prom, pred_prom, prob_prom)

```




```{r}

set.seed(10)
rf_model <- randomForest(is_promoted ~ ., trn.data[,-1], ntree = 1000,
                         do.trace = 50, importance = T)

```




```{r}

varImp(rf_model)

varImpPlot(rf_model)

```





```{r}

rf_pred <- predict(rf_model, newdata = val.data, type = 'prob')

summary(rf_pred)

```




```{r}

rf_prom <- as.factor(ifelse(rf_pred[,2] > 0.05, '1', '0'))

confusionMatrix(rf_prom, act_prom, positive = '1')


```





```{r}

# Balancing the classes by SMOTE
require(smotefamily)

set.seed(100)
trn.smote <- SMOTE(trn.data[,-c(1,32)], trn.data$is_promoted)

# Classes are well balanced as compared to orig. data.

```


```{r}
trn.data <- trn.smote$data

str(trn.data)

```



```{r}
names(trn.data)[33] <- 'is_promoted'

```



```{r}
table(trn.data$is_promoted)


```




```{r}

sapply(trn.data, n_distinct)

```



```{r}

trn.data[,c(1:18, 21:26, 28:29)] <- sapply(trn.data[,c(1:18, 21:26, 28:29)], round)

```




```{r}
require(class)
# KNN on balanced classes

knn_pred <- knn(train = select(trn.data, -is_promoted), 
                cl = trn.data$is_promoted, 
                test = select(val.data, -c(employee_id, is_promoted)), 
                k = 25)

confusionMatrix(knn_pred, act_prom, positive = '1')


```



```{r}
dt_model <- rpart(is_promoted ~ ., data = trn.data)

prp(dt_model)

```


```{r}

dt_pred <- predict(dt_model, newdata = val.data)

dt_prom <- as.factor(ifelse(dt_pred[,2] > 0.7, "1", "0"))

confusionMatrix(dt_prom, act_prom, positive = '1')


```


```{r}
rm(dt_model, dt_pred, model_3, rf_pred, trn.smote, dt_prom, knn_prom, rf_prom)
```


```{r}

# Building SVM model

require(e1071)

?e1071::svm

trn.data$is_promoted <- as.factor(trn.data$is_promoted)

# Linear
svm_lnr <- svm(is_promoted ~ ., data = trn.data, kernel = "linear", scale = F)

```




```{r}

pred_prom <- predict(svm_lnr, newdata = val.data)

confusionMatrix(pred_prom, act_prom, positive = "1")

```



```{r}
trn.data$is_promoted <- as.factor(trn.data$is_promoted)

svm_rad <- svm(is_promoted ~ ., data = trn.data, kernel = "radial", scale = F)

```

```{r}

pred_prom <- predict(svm_rad, newdata = val.data)

confusionMatrix(pred_prom, act_prom, positive = "1")


```




Tuning hyper-parameters by grid-search



```{r}
svm_tune <- tune(svm, train.x = select(trn.data, -is_promoted), 
                 train.y = trn.data$is_promoted, kernel="radial", 
                 ranges = list(cost = 10^(-2:2), gamma = 10^(-2:2) ))

#tune() gives us the tuned parameters, C and gamma

df <- summary(svm_tune)
df

```





```{r}
plot(df$performances$error, type = 'o', col='red')

df$best.parameters

svm_model_after_tune <- svm(Species ~ ., data=train,
                            method = "C-classification", 
                            kernel="radial", cost=10, gamma=0.1)

table(predict(svm_model_after_tune, test[,-5]), test[,5])


```


*******************************************************************************************

