


```{r}
require(dplyr)
require(readr)
require(caTools)
require(caret)
require(ggplot2)

```

# load dataset
```{r}
data <- read_csv('E:/coaching/Data Science/R/dataset/Adult(KNN)/DataSet/adult.csv',show_col_types = FALSE)

glimpse(data)

```


# missing data
```{r}
colSums(is.na(data))

```

# but we can see here ? , it seems to be missing values

# we should replace ? with NAs

```{r}
data <- read_csv('E:/coaching/Data Science/R/dataset/Adult(KNN)/DataSet/adult.csv',show_col_types = FALSE, na = c("?","NA"))

```

```{r}
#mismatch
data <- mutate_if(data, is.character, tolower)

# duplicates
data <- distinct(data)

```



# univariate analysis and treat missing values also / CATEGORICAL COLUMNS
```{r}
sapply(data, n_distinct)

```


# WORK CLASS

```{r}
table(data$workclass)

```

# we should combine the categories for our convienence

# never_worked or without_pay = jobless
# local gov or state gov = government
# self_emp_inc or not_inc = selfemployed

```{r}
data$workclass[data$workclass == "without-pay" | data$workclass == "never-worked"] <- "jobless"


data$workclass[data$workclass == "local-gov" | data$workclass == "state-gov"|
                 data$workclass=='federal-gov' | data$workclass == "government"] <- "government"


data$workclass[data$workclass == "self-emp-inc" | data$workclass == "self-emp-not-inc"] <- "self-employed"

```



```{r}
table(data$workclass)
sum(is.na(data$workclass))

```

# we make the new category of unknown for missing values


```{r}
data$workclass[is.na(data$workclass)] <- "unknown"

table(data$workclass)

```




# NATIVE COUNTRY
```{r}

summary(as.factor( data$`native-country`))

```

# we also combine the cities and country to their respective country

# cambodia,china,hong,india,iran,japan,laos,phillippines,taiwan,thailand,vietnam - asia

# canada,cuba,dominanican replubic,el-salvador,guatemala,haiti,honduras,jamaica,mexico
# ,nicaragua,outlying-us,puerto,trinada, unitedstaes - north america

# columbia,ecauado,peru <- south_america

# england,france,germany,greece,holand-netherland,hungary,ireland,italy,poland,portugal,
#  scotland,yugoslavia - europe

# south,na - others



```{r}
data$`native-country`[data$`native-country`!="united-states"] <- "outside-united-states"

```





```{r}
ggplot(data, aes(x = `native-country`, fill = income)) +geom_bar(position = 'fill')
```


```{r}
data$`native-country`[is.na(data$`native-country`)]  <- "united-states"

```



```{r}
summary(as.factor( data$`native-country`))

```

```{r}
colSums(is.na(data))
```


```{r}
# OCCUPATION

table(data$occupation)

```

```{r}
data$occupation[data$occupation=="other-service"|data$occupation=="priv-house-serv"|
                  data$occupation=="protective-serv"] <- "services"

data$occupation[data$occupation=="craft-repair"|data$occupation=="farming-fishing"|
                  data$occupation=="handlers-cleaners"|data$occupation=="machine-op-inspct"|
                  data$occupation=="transport-moving"] <- "blue_collar"

data$occupation[data$occupation=="exec-managerial"|data$occupation=="prof-specialty"|
                  data$occupation=="tech-support"] <- "white_collar"

```



```{r}
data$occupation[is.na(data$occupation)] <- "unknown"

sum(is.na(data$occupation))

table(data$occupation)

```

  # NOW MISSING VALUES GONE

# EDUCATION

```{r}
table(data$education)

```

# combine categories
```{r}

data$education[data$education=="10th"|data$education=="11th"|data$education=="12th"|
                 data$education=="1st-4th"|data$education=="5th-6th"|data$education=="7th-8th"|
                 data$education=="9th"|data$education=="preschool"] <- "school"

data$education[data$education=="assoc-acdm"|data$education=="assoc-voc"|data$education=="prof-school"] <- 
  "associate_and_voc"


data$education[data$education=="bachelors"|data$education=="hs-grad"] <- "graduates"

```


```{r}
table(data$education)


colSums(is.na(data))

```


# EDUCATIONAL-NUM

```{r}
table(data$`educational-num`)

```

# MARITAL STATUS

```{r}


table(data$`marital-status`)

data$`marital-status`[data$`marital-status`=="married-af-spouse"|data$`marital-status`=="married-civ-spouse"|
                        data$`marital-status`=="married-spouse-absent"] <- "married"


data$`marital-status`[data$`marital-status`=="never-married"] <- "unmarried"


data$`marital-status`[data$`marital-status`=="divorced"|data$`marital-status`=="separated"|
                        data$`marital-status`=="widowed"] <- "others"

table(data$`marital-status`)

```


# RELATIONSHIP

table(data$relationship)

# RACE

table(data$race)


# GENDER

table(data$gender)

                               # CONTINOUS VARIABLES


# AGE

```{r}
plot(quantile(data$age,seq(0,1,0.01)))

data$age[data$age>74] <- 74

```


# FNLWGHT

```{r}
summary(data$fnlwgt)
plot(quantile(data$fnlwgt,seq(0,1,0.01)))

data$fnlwgt[data$fnlwgt > 509595.46] <- 509595.46

```




# capital gain

```{r}
summary(data$`capital-gain`)

plot(quantile(data$`capital-gain`,seq(0,1,0.01)))

quantile(data$`capital-gain`,seq(0,1,0.01))

data$`capital-gain`[data$`capital-gain` > 15024] <- 15024

```


```{r}
# capital loss
summary(data$`capital-loss`)
table(data$`capital-loss`,data$income)

plot(quantile(data$`capital-loss`,seq(0,1,0.01)))

quantile(data$`capital-loss`,seq(0,1,0.01))


```


# hours - per-week
```{r}

summary(data$`hours-per-week`)
plot(quantile(data$`hours-per-week`,seq(0,1,0.01)))

```


```{r}
data$`hours-per-week`[data$`hours-per-week`>80] <- 80

```


# BIVARIATE ANALYSIS AND VISULASIATIONS

# AGE AND INCOME

ggplot(data,aes(age))+geom_histogram(aes(fill = as.factor(data$income)),color="black",binwidth = 1)

# AGE AND GENDER
ggplot(data,aes(age))+geom_histogram(aes(fill = as.factor(data$gender)),color="black",binwidth = 1)

 # workclass distribution
ggplot(data,aes(x = workclass))+geom_bar(fill="skyblue")

# more adults belongs to private_jobs
 # WORKCLASS AND INCOME
ggplot(data,aes(x=workclass,y=as.factor(income)))+geom_col()


# FNLWGHT AND INCOME

ggplot(data,aes(x=fnlwgt))+geom_histogram(fill=as.factor(data$income),color="black",binwidth = 1)


#hours per-week

ggplot(data,aes(x="hours-per-week")+geom_histogram(binwidth = 1)
      

#  marital status
   # disribution of marital staus
ggplot(data,aes(x = `marital-status`))+geom_bar(fill="skyblue")

   # marital staus vs income
ggplot(data,aes(x=income,fill=`marital-status`))+geom_bar(position = "dodge")

```{r}
df1 <- data.frame(table(data$income,data$`marital-status`))

names(df1) <- c("income","status","count")


ggplot(df1,aes(x=status,y=count,fill=income))+geom_bar(stat="identity")

```



# OCCUPATION
    # distribution of occupation

ggplot(data,aes(x=occupation))+geom_bar()

    # occupation vs income

 ggplot(data,aes(x=income,fill=occupation))+geom_bar(position = "dodge")

df2 <- data.frame(table(data$income,data$occupation))

names(df2) <- c("income","occu","count")


ggplot(df2,aes(x=occu,y=count,fill=income))+geom_bar(stat="identity")


# GENDER

ggplot(data,aes(x=gender))+geom_bar()

   # gender vs income

ggplot(data,aes(x=income,fill=gender))+geom_bar(position = "dodge")

df3 <- data.frame(table(data$income,data$gender))

names(df3) <- c("income","gender","count")


ggplot(df3,aes(x=gender,y=count,fill=income))+geom_bar(stat="identity")


# RACE


ggplot(data,aes(x=race))+geom_bar()

# race vs income

ggplot(data,aes(x=income,fill=race))+geom_bar(position = "dodge")

df4 <- data.frame(table(data$income,data$race))

names(df4) <- c("income","race","count")


ggplot(df4,aes(x=race,y=count,fill=income))+geom_bar(stat="identity")

# EDUCATION


ggplot(data,aes(x=education))+geom_bar()

# education vs income

ggplot(data,aes(x=income,fill=education))+geom_bar(position = "dodge")

df5 <- data.frame(table(data$income,data$education))

names(df5) <- c("income","education","count")


ggplot(df5,aes(x=education,y=count,fill=income))+geom_bar(stat="identity")


# EDUCATIONAL-NUM
colnames(data)
ggplot(data,aes(x=`educational-num`))+geom_bar()

# educationAL-NUM vs income

ggplot(data,aes(x=income,fill=`educational-num`))+geom_bar(position = "dodge")

df6 <- data.frame(table(data$income,data$`educational-num`))

names(df6) <- c("income","education","count")


ggplot(df6,aes(x=education,y=count,fill=income))+geom_bar(stat="identity")

# RELATIONSHIP

ggplot(data,aes(x=relationship))+geom_bar()

# relationship vs income

ggplot(data,aes(x=income,fill=relationship))+geom_bar(position = "dodge")

df7 <- data.frame(table(data$income,data$relationship))

names(df7) <- c("income","relationship","count")


ggplot(df7,aes(x=relationship,y=count,fill=income))+geom_bar(stat="identity")

# NATIVE COUNTRY

ggplot(data,aes(x=`native-country`))+geom_bar()

# `native-country` vs income

ggplot(data,aes(x=income,fill=`native-country`))+geom_bar(position = "dodge")

df8 <- data.frame(table(data$income,data$`native-country`))

names(df8) <- c("income","country","count")


ggplot(df8,aes(x=country,y=count,fill=income))+geom_bar(stat="identity")


# DROPPING VARIABLES

data$fnlwgt <- NULL    
data$education <- NULL    bcz educational_num is given
data$relationship <- NULL   no use in prediction


# SOME OTHER ANAMOLIES

```{r}
data$income <- ifelse(data$income == "<=50k",0, 1)
str(data)

```



# feature scaling

```{r}
data[,c(1,3,5, 11:13)] <- sapply(data[,c(1,3,5, 11:13)], scale)

```

# MODEL BUILDING


```{r}
summary(data)
```


```{r}
require(dummies)
final_data <- dummy.data.frame(as.data.frame(data))
```



# split data in to train and test

```{r}

set.seed(111)
i <- sample.split(final_data$income , SplitRatio = 0.75)

train <- final_data[i,]

test <- final_data[!i,]


```


```{r}


train_labels <- as.factor(train$income)

test_labels <- as.factor(test$income)


```


                                       # KNN

```{r}
require(class)

help(knn)

```

```{r}

knn_labels <- knn(train = train[,-43], cl = train_labels, test = test[,-43], k = 5)


```


# CONFUSION MATRIX 
```{r}
confusionMatrix(knn_labels, test_labels, positive = "1")

```

```{r}
knn_labels <- knn(train = train[,-43], cl = train_labels, test = test[,-43], k = 10)

confusionMatrix(knn_labels, test_labels, positive = "1")

```
```{r}
cm <- confusionMatrix(knn_labels, test_labels, positive = "1")
```

```{r}
k_optm <- data.frame(k=rep(0,50), acc =rep(0,50), sens=rep(0,50), spec=rep(0,50))
```


```{r}
k_optm <- data.frame(k=rep(0,50), acc =rep(0,50), sens=rep(0,50), spec=rep(0,50))

for (i in 2:50) {
  knn_labels <- knn(train = train[,-43], cl = train_labels, test = test[,-43], k = i)
  cm <- confusionMatrix(knn_labels, test_labels, positive = "1")
  
  k_optm$k = 1
  k_optm$acc[i] <- cm$overall[1]
  k_optm$sens[i] <- cm$byClass[1]
  k_optm$spec[i] <- cm$byClass[2]
  }
```

```{r}
k_optm[which.max(k_optm$acc),]
```
```{r}
k_optm[which.max(k_optm$sens),]
```
```{r}
k_optm[which.max(k_optm$spec),]
```

